---
title: "Introduction to Bayesian kernel machine regression"
author: "Jennifer F. Bobb"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Bayesian kernel machine regression}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r options, echo=FALSE}
## if the current working directory is the directory where this file is located rather than the project directory, set the working directory to the project directory
if(grepl("vignettes", getwd())) knitr::opts_knit$set(root.dir = paste(getwd(), "..", sep = "/"))
#knitr::opts_chunk$set(fig.width = 7, fig.height = 5, message = FALSE)
```

```{r setup, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
devtools::document()
devtools::load_all()
```

## Simulate some data

```{r simdata}
set.seed(5)
dat <- SimData1(n = 100, M = 5, sigsq.true = 0.5, beta.true = 2)
y <- dat$y
expos <- dat$expos
covar <- dat$covar
```

## Conduct some exploratory data analysis

```{r boxplots, fig.height=4, fig.width=6}
boxplot(expos)
```

## Prior distributions for the $r_m$ parameters

```{r investigate prior}
fits <- InvestigatePrior(y = y, expos = expos, covar = covar)
```

Here we plot the estimated univariate exposure-response relationships for `r_m` fixed to different values.
```{r investigate prior 2, fig.height=7, fig.width=7}
PlotPriorFits(y = y, expos = expos, covar = covar, fits = fits, 
              which.q = c(1,3,4,6), which.expos = 1:3)
```

## Fit BKMR 

### Using Stan software

```{r fit stan, eval=FALSE}
set.seed(123)
runtime <- system.time({
    fitstan <- StanBKMR(y = y, expos = expos, covar = covar, iter = 300, chains = 4)
})
```

Because the above code took 51 minutes to run on my computer, I have saved the output as the dataset `fitstan` that is provided with this package. 
```{r load}
print(fitstan, par = c("beta", "r", "sigma_sq", "h[1]", "h[2]", "h[100]"))
```

We see that the estimated $\beta$ and $\sigma^2$ are close to their true values.

### Using R code

```{r fit orig, eval=FALSE}
set.seed(456)
runtime <- system.time(
    fitbkmr <- kmbayes(y = y, expos = expos, covar = covar, iter = 10000, quiet = FALSE)
)
```

```{r load orig}
fitbkmr
```

## Summarize BKMR output

### Plot the exposure-response function









































