---
title: "Introduction to Bayesian kernel machine regression"
author: "Jennifer F. Bobb"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Bayesian kernel machine regression}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r options, echo=FALSE, message=FALSE}
## if the current working directory is the directory where this file is located rather than the project directory, set the working directory to the project directory
if(grepl("vignettes", getwd())) knitr::opts_knit$set(root.dir = paste(getwd(), "..", sep = "/"))
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, message = FALSE)
```

```{r setup, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
##library(bkmr)
devtools::document()
##devtools::load_all()
```

## Simulate some data

```{r simdata}
set.seed(111)
dat <- SimData(n = 50, M = 4, sigsq.true = 0.5, beta.true = 2, HFun = HFun3)
y <- dat$y
expos <- dat$expos
covar <- dat$covar
```

## Conduct some exploratory data analysis

```{r cormat}
PlotCorr(expos, print.vals = TRUE)
```

## Prior distributions for the $r_m$ parameters

```{r investigate prior}
priorfits <- InvestigatePrior(y = y, expos = expos, covar = covar)
```

Here we plot the estimated univariate exposure-response relationships for `r_m` fixed to different values.
```{r investigate prior 2, fig.height=6, fig.width=6}
PlotPriorFits(y = y, expos = expos, covar = covar, 
              fits = priorfits, which.q = c(1,3,4,6))
```

## Fit BKMR 

```{r fit orig}
set.seed(111)
runtime <- system.time(
    fitkm <- kmbayes(y = y, expos = expos, covar = covar, iter = 10000, quiet = TRUE, modsel = TRUE)
)
```

## Investigate model convergence

Let's visually inspect the trace plots, showing how various parameter values change as the sampler runs.

```{r trace plots, fig.height=2.5, fig.width=6}
TracePlot(fit = fitkm, par = "beta")
TracePlot(fit = fitkm, par = "sigsq.eps")
TracePlot(fit = fitkm, par = "r", comp = 1)
TracePlot(fit = fitkm, par = "h", comp = 20)
```

## Summarize output

### Summary statistics of the exposure-response function

Compute the overall effect of the mixture, by comparing when all exposures are at a particular percentile to when all are at the 50th percentile.

```{r overall}
risks.overall.approx <- OverallRiskSummaries(fit = fitkm, y = y, expos = expos, covar = covar)
```

```{r plot overall}
library(ggplot2)
ggplot(risks.overall.approx, aes(quantile, est, ymin = est - 1.96*se, ymax = est + 1.96*se)) + 
    geom_pointrange()
```

Compute summary statistics summarizing the single pollutant health effects; compare risk when a single pollutant is at the 75th versus 25th percentile, when all of the other exposures are fixed at a particular percentile. We refer to this as the single-pollutant health risks.

```{r single pol}
risks.singpol <- SingPolRiskSummaries(fit = fitkm, y = y, expos = expos, covar = covar)
```

```{r plot single pol}
ggplot(risks.singpol, aes(exposure, est, ymin = est - 1.96*se, ymax = est + 1.96*se, 
                          col = q.fixed)) + 
    geom_pointrange(position = position_dodge(width = 0.75)) + coord_flip()
```

Compute summary statistics summarizing interaction; compare the single-pollutant health risks when all of the other pollutants are fixed to their 75th percentile to when all of the other pollutants are fixed to their 25th percentile.

```{r int}
risks.int <- SingPolIntSummaries(fit = fitkm, y = y, expos = expos, covar = covar)
```

If the 95% credible interval does not cover zero for a particular pollutant, this suggests that that pollutant may interact with one of the other pollutants.
```{r plot int}
ggplot(risks.int, aes(exposure, est, ymin = est - 1.96*se, ymax = est + 1.96*se)) + 
    geom_hline(y_intercept = 0, col = "brown", lty = 2) +
    geom_pointrange()
```

### Plot the exposure-response function

Single-pollutant exposure-response function for other exposures fixed at their 50th percentile.

```{r expos-resp}
expos.resp.univar <- ExposureResponseUnivar(fit = fitkm, y = y, expos = expos, covar = covar)
```

```{r plot expos-resp, fig.height=4, fig.width=6}
ggplot(expos.resp.univar, aes(z, est, ymin = est - 1.96*se, ymax = est + 1.96*se)) + 
    geom_smooth(stat = "identity") + 
    facet_wrap(~ exposure)
```

Two-pollutant exposure-response function for other exposures fixed at their 50th percentile.

```{r expos-resp2, message=FALSE}
expos.resp.bivar <- ExposureResponseBivar(fit = fitkm, y = y, expos = expos, covar = covar, 
                                          min.plot.dist = 1)
```

Sometimes the image plots are hard to visualize:
```{r plot expos-resp2, fig.height=5, fig.width=6.5}
ggplot(expos.resp.bivar, aes(z1, z2)) + 
    geom_raster(aes(fill = est)) + 
    facet_grid(exposure1 ~ exposure2)
```

Another option for investigating bivariate exposure-response relations is to plot the exposure-response function of a single pollutant for the second pollutant fixed at various quantiles.
```{r expos-resp2 opt2, message=FALSE}
expos.resp.bivar.levels <- ExposureResponseBivarLevels(expos.resp.df = expos.resp.bivar, 
                                                       expos = expos, qs = c(0.25, 0.5, 0.75))
```

```{r plot expos-resp2 opt2, fig.height=5, fig.width=6.5, warning=FALSE}
ggplot(expos.resp.bivar.levels, aes(z1, est)) + 
    geom_smooth(aes(col = quantile), stat = "identity") + 
    facet_grid(exposure1 ~ exposure2) +
    ggtitle("f(z1 | quantiles of z2)")
```

































